SDGs

### 2022-03-02 Trilemma_clustering.Rmd
data explorations for SDGs datasets

# ggsave(
#     plot = p, path = "figures/", file = "first_diff.png", device = "png",
#     width = 6, height = 3, dpi = 300, bg = "white")
p + theme_light(base_size = 20)
```
```{r fig.width=3, fig.height=3}
c1 <- df_dat |> 
    ggplot(aes(EFconsPerCap, gini)) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = 1.68, linetype = 2, color = "gray50") +
    geom_density2d() +
    geom_point(aes(color = gni_log)) +
    geom_path(aes(group = country), size = 0.15) +
    scico::scale_color_scico("GNI [log]") +
    labs(y = "Gini", x = "Ecological footprint") +
    theme_light(base_size = 7) +
    theme(legend.position = "top", panel.grid = element_blank())

c1 + theme_light(base_size = 18)
```
```{r fig.width=3, fig.height=3}
c2 <- df_dat |> 
    ggplot(aes(gni2, gini)) +
    geom_density2d() +
    geom_point(aes(color = EFconsPerCap)) +
    geom_path(aes(group = country), size = 0.15) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = (12546), linetype = 2, color = "gray50") +
    scico::scale_color_scico("Ecological footprint",palette = "berlin") +
    labs(x= "GNI [log]", y = "Gini") +
    scale_x_log10() +
    theme_light(base_size = 7) +
    theme(legend.position = "top", panel.grid = element_blank())

c2 + theme_light(base_size = 7)
```
```{r fig.width=3, fig.height=3}
c3 <- df_dat |> 
    ggplot(aes(gni2, EFconsPerCap)) +
    geom_density2d() +
    geom_point(aes(color = gini)) +
    geom_path(aes(group = country), size = 0.15) +
    scico::scale_color_scico("Gini") +
    labs(x = "GNI", y = "Ecological footprint") +
    scale_x_log10()+
    geom_hline(yintercept = 1.68, linetype = 2, color = "gray50") +
    geom_vline(xintercept = (12546), linetype = 2, color = "gray50") +
    theme_light(base_size = 7) +
    theme(legend.position = "top", panel.grid = element_blank())

@@

```{r fig.width=6, fig.height=2}
df_summary <- df_dat |> 
    as_tibble() |> 
    ungroup() |> 
    group_by(country) |> 
    dplyr::summarize(sd_magnitude = sd(magni, na.rm = TRUE),
              sd_theta = sd(theta, na.rm = TRUE),
              sd_gamma = sd(gamma, na.rm = TRUE), 
              EF = mean(EFconsPerCap, na.rm = TRUE),
              Gini = mean(gini, na.rm = TRUE),
              GNI = mean(gni_log, na.rm = TRUE),
              sd_EF = sd(EFconsPerCap, na.rm = TRUE),
              sd_Gini = sd(gini, na.rm = TRUE),
              sd_GNI = sd(gni_log, na.rm = TRUE),
              magnitude_m = mean(magni, na.rm = TRUE), 
              theta_m = mean(theta, na.rm = TRUE),
              gamma_m = mean(gamma, na.rm = TRUE),
              magnitude_a = sum(magni, na.rm = TRUE), 
              theta_a = sum(theta, na.rm = TRUE),
              gamma_a = sum(gamma, na.rm = TRUE))

@@

clust_num <- NbClust::NbClust( 
    data = df_summary %>% as.data.frame() %>% 
        select(where(is.numeric)),
    #distance = "maximum",
    min.nc = 2, max.nc = 7, 
    method = m, 
    index = 'all')
```

@@

```{r}
## Stability & Internal validation
stab <- clValid::clValid(
    obj = df_summary %>% as.data.frame() %>% 
        select(where(is.numeric)) %>%
        as.matrix(),
    nClust = c(2:7),
    clMethods = c("hierarchical", "kmeans",  #"som", "model", "diana", "fanny",
                 "sota", "pam", "clara", "agnes"),
    validation = c('stability', "internal"),
    #metric = "manhattan",
    method = "ward.D2",
    verbose = FALSE
) ## Hierarchical is the optimal method

@@

```{r}
stab@clusterObjs$pam$`2`$clustering
```

```{r fig.width=7, fig.height=8}
df_summary |> 
    mutate(clus = stab@clusterObjs$pam$`2`$clustering) |> 
    mutate(country = as_factor(country)) |> 
    mutate(country = fct_reorder2(country, magnitude, Gini)) |> 
    pivot_longer(cols = sd_magnitude:gamma, names_to = "vars", values_to = "vals") |> 
    ggplot(aes(vals, country)) +
    geom_col(aes(fill = as.factor(clus))) +
    facet_grid(clus~vars, scales = "free") +
    theme_light(base_size = 10)
```

```{r}
df_summary |>
    select(-starts_with("country")) |> 
    GGally::ggpairs()
```

@@

pc1 <- ggbiplot::ggbiplot(pc, groups = as.character(stab@clusterObjs$pam$`2`$clustering),
         ellipse = TRUE, labels = df_summary$country_code, labels.size = 2.5,
         varname.size = 4) +
    scale_color_manual("Groups", values = c("orange", "purple")) +
    theme_light(base_size = 18) +
    theme(legend.position = c(0.2, 0.2), panel.grid = element_blank())

pc1
```

Note to self: It does not make sense to plot the average distance traversed in the space per country. Take Luxemburg as example, it started and ended verly close by. The absolute distance is what I'm calculating with the euclidian distance on that plane, then the first-diff values are just increases or decreases with respect to initial conditions. Averages are then misleading, because on average Luxemburg moved a lot (in circles).

Averages are worth calculating for the parameter space, the vector field of the cube. But not for individual countries, or particules moving within the cube. Be aware of that when presenting and interpreting the results. 
```{r fig.height=3,fig.width=3}
a1 <- df_dat |> 
    as_tibble() |> 
    group_by(country) |> 
    dplyr::summarize(
        EF = sum(d_EF, na.rm = TRUE),
        Gini = sum(d_gini, na.rm = TRUE),
        GNI = sum(d_gni, na.rm = TRUE)
    ) |> 
    mutate(Group = as.character(stab@clusterObjs$pam$`2`$clustering)) |> 
    left_join(
        dat |> select(country, country_code) |> 
            filter(!is.na(country_code)) |> unique()) |> 
    ggplot(aes(EF, Gini)) +
    geom_segment(
        aes(x = 0, y = 0, xend = EF, yend = Gini, color = as.factor(Group)),
        arrow = arrow(length = unit(0.15, "cm")), size = 0.15) +
    geom_text(
        aes(x = EF *1.1, y = Gini * 1.1, label = country_code, colour = Group), size = 2) +
    scale_color_manual("Groups", values = c("orange", "purple")) +
    labs(x = "Ecological Footprint", y = "Gini") +
    theme_light(base_size = 6) +
    theme(panel.grid = element_blank(), legend.position = c(0.8,0.2))

@@

c1 + annotation_custom(
        grob = ggplotGrob(a1),
        xmin = 10, xmax = 18, ymin = 40, ymax = 60
    ) 
```

```{r fig.width=2, fig.height=2}
a2 <- df_dat |> 
    as_tibble() |> 
    group_by(country) |> 
    dplyr::summarize(
        EF = sum(d_EF, na.rm = TRUE),
        Gini = sum(d_gini, na.rm = TRUE),
        GNI = sum(d_gni, na.rm = TRUE)
    ) |> 
    mutate(Group = as.character(stab@clusterObjs$pam$`2`$clustering)) |> 
    left_join(
        dat |> select(country, country_code) |> 
            filter(!is.na(country_code)) |> unique()) |> 
    ggplot(aes(GNI, Gini)) +
    geom_segment(
        aes(x = 0, y = 0, xend = GNI, yend = Gini, 
            color = as.factor(Group)), 
        arrow = arrow(length = unit(0.15, "cm")), size = 0.15) +
    geom_text(aes(x = GNI * 1.1, y = Gini *1.1, 
                  label = country_code, colour = Group), size = 2) +
    scale_color_manual("Groups", values = c("orange", "purple")) +
    theme_light(base_size = 6) +
    theme(panel.grid = element_blank(), legend.position = c(0.8,0.2))

c2 + annotation_custom(
        grob = ggplotGrob(a2),
        xmin = 1e+01, xmax = 1e+03, ymin = 40, ymax = 60
    )

a2
```

>>>>

# ggsave(
#     plot = p, path = "figures/", file = "first_diff.png", device = "png",
#     width = 6, height = 3, dpi = 300, bg = "white")
p + theme_light(base_size = 20)
```
```{r fig.width=3, fig.height=3}
c1 <- df_dat |> 
    ggplot(aes(EFconsPerCap, gini)) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = 1.68, linetype = 2, color = "gray50") +
    geom_density2d(size = 0.25, color = "#446B9C") +
    geom_point(aes(color = gni_log), size = 0.5) +
    geom_path(aes(group = country), size = 0.1) +
    scico::scale_color_scico("GNI [log]") +
    labs(y = "Gini", x = "Ecological footprint") +
    theme_light(base_size = 6) +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(7, "mm"), legend.key.height = unit(2,"mm"))

c1 + theme_light(base_size = 18)
```
```{r fig.width=3, fig.height=3}
c2 <- df_dat |> 
    ggplot(aes(gni2, gini)) +
    geom_density2d(size = 0.25, color = "#446B9C") +
    geom_point(aes(color = EFconsPerCap), size = 0.5) +
    geom_path(aes(group = country), size = 0.1) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = (12546), linetype = 2, color = "gray50") +
    scico::scale_color_scico("Ecological footprint",palette = "berlin") +
    labs(x= "GNI [log]", y = "Gini") +
    scale_x_log10() +
    theme_light(base_size = 6) +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(7, "mm"), legend.key.height = unit(3,"mm"))

c2 + theme_light(base_size = 7)
```
```{r fig.width=3, fig.height=3}
c3 <- df_dat |> 
    ggplot(aes(gni2, EFconsPerCap)) +
    geom_density2d(size = 0.25, color = "#446B9C") +
    geom_point(aes(color = gini), size =0.5) +
    geom_path(aes(group = country), size = 0.1) +
    scico::scale_color_scico("Gini") +
    labs(x = "GNI", y = "Ecological footprint") +
    scale_x_log10()+
    geom_hline(yintercept = 1.68, linetype = 2, color = "gray50") +
    geom_vline(xintercept = (12546), linetype = 2, color = "gray50") +
    theme_light(base_size = 6) +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(7, "mm"), legend.key.height = unit(3,"mm"))

@@

```{r fig.width=6, fig.height=2}
df_summary <- df_dat |> 
    as_tibble() |> 
    ungroup() |> 
    group_by(country) |> 
    dplyr::summarize(sd_magnitude = sd(magni, na.rm = TRUE),
              sd_theta = sd(theta, na.rm = TRUE),
              sd_gamma = sd(gamma, na.rm = TRUE), 
              EF = mean(EFconsPerCap, na.rm = TRUE),
              Gini = mean(gini, na.rm = TRUE),
              GNI = mean(gni_log, na.rm = TRUE),
              sd_EF = sd(EFconsPerCap, na.rm = TRUE),
              sd_Gini = sd(gini, na.rm = TRUE),
              sd_GNI = sd(gni_log, na.rm = TRUE),
              #magnitude_m = mean(magni, na.rm = TRUE),
              #theta_m = mean(theta, na.rm = TRUE),
              #gamma_m = mean(gamma, na.rm = TRUE),
              magnitude_a = sum(magni, na.rm = TRUE),
              theta_a = sum(theta, na.rm = TRUE),
              gamma_a = sum(gamma, na.rm = TRUE)
              )

@@

clust_num <- NbClust::NbClust( 
    data = (df_summary %>% as.data.frame() %>% 
        select(where(is.numeric)) ),
    #distance = "maximum",
    min.nc = 2, max.nc = 7, 
    method = m, 
    index = 'all')
```

@@

```{r warning = FALSE, message=FALSE}

library(clValid)
library(mclust)
## Stability & Internal validation
stab <- clValid::clValid(
    obj = df_summary %>% as.data.frame() %>% 
        select(where(is.numeric)) %>%
        as.matrix(),
    nClust = 3,
    clMethods = c(
        "hierarchical", "kmeans",  "som", 
        "model", "diana", "sota", "pam", "clara", "agnes"),
    validation = c('stability', "internal"),
    #metric = "manhattan",
    method = "ward",
    verbose = FALSE
) ## Hierarchical is the optimal method

@@

```{r}
# stab@clusterObjs$pam$`2`$clustering

stab@clusterObjs$kmeans$`3`$cluster

stab@clusterObjs$diana |> plot()
```

```{r fig.width=7, fig.height=8}
df_summary |> 
    mutate(clus = stab@clusterObjs$kmeans$`3`$cluster ) |> 
    mutate(country = as_factor(country)) |> 
    mutate(country = fct_reorder2(country, magnitude_a, Gini)) |> 
    pivot_longer(
        cols = sd_magnitude:gamma_a, names_to = "vars", 
        values_to = "vals") |> 
    ggplot(aes(vals, country)) +
    geom_col(aes(fill = as.factor(clus))) +
    facet_grid(clus~vars, scales = "free") +
    theme_light(base_size = 10)
```

```{r message = FALSE, warning = FALSE, fig.width=5, fig.height=5}
df_summary |>
    select(-starts_with("country")) |> 
    GGally::ggpairs()
```

@@

pc1 <- ggbiplot::ggbiplot(
    pc, groups = as.character(stab@clusterObjs$kmeans$`3`$cluster),
    ellipse = TRUE, labels = df_summary$country_code, labels.size = 2.5,
    varname.size = 3) +
    scale_color_manual("Groups", values = c("#73B3A3","#FEA621","#5BA4CA")) +
    theme_light(base_size = 6) +
    theme(legend.position = c(0.2, 0.2), panel.grid = element_blank(),
          legend.text = element_text(size = 6), legend.title = element_text(size = 6),
          legend.key.width = unit(3,"mm"), legend.key.height = unit(2, "mm"))

pc1

ggsave(
    filename = "pca_trilema.png", path = "figures/", device = "png",
    width = 4, height = 4, dpi = 300, bg = "white", 
    plot = pc1
)

Note to self: It does not make sense to plot the average distance traversed in the space per country. Take Luxemburg as example, it started and ended very close by. The absolute distance is what I'm calculating with the euclidian distance on that plane, then the first-diff values are just increases or decreases with respect to initial conditions. Averages are then misleading, because on average Luxembourg moved a lot (in circles).

Averages are worth calculating for the parameter space, the vector field of the cube. But not for individual countries, or particules moving within the cube. Be aware of that when presenting and interpreting the results. 
```{r fig.height=3,fig.width=3}
a1 <- df_dat |> 
    as_tibble() |> 
    group_by(country) |> 
    dplyr::summarize(
        EF = sum(d_EF, na.rm = TRUE),
        Gini = sum(d_gini, na.rm = TRUE),
        GNI = sum(d_gni, na.rm = TRUE)
    ) |> 
    mutate(Group = as.character(stab@clusterObjs$kmeans$`3`$cluster)) |> 
    left_join(
        dat |> select(country, country_code) |> 
            filter(!is.na(country_code)) |> unique()) |> 
    ggplot(aes(EF, Gini)) +
    geom_segment(
        aes(x = 0, y = 0, xend = EF, yend = Gini, color = as.factor(Group)),
        arrow = arrow(length = unit(0.1, "cm")), show.legend = FALSE, size = 0.15) +
    geom_text(
        aes(x = EF *1.1, y = Gini * 1.1, label = country_code, colour = Group), 
        size = 1, show.legend = FALSE) +
    scale_color_manual("Groups", values = c("#73B3A3","#FEA621","#5BA4CA")) +
    labs(x = "Ecological Footprint", y = "Gini") +
    theme_light(base_size = 6) +
    theme(panel.grid = element_blank(), legend.position = c(0.8,0.2), 
          legend.text = element_text(size = 4), legend.title = element_text(size =4),
          legend.key.width = unit(2,"mm"), legend.key.height = unit(2, "mm"))

@@

f1 <- c1 + annotation_custom(
        grob = ggplotGrob(a1),
        xmin = 10, xmax = 18, ymin = 40, ymax = 60
    )

# ggsave(
#     filename = "Gini_EF.png", path = "figures/", device = "png",
#     width = 4, height = 4, dpi = 300, bg = "white", 
#     plot = f1
# )

 ggplotly(c1, tooltip=c("country", "EFconsPerCap", "gini", "gni_log"))

```{r fig.width=3, fig.height=3}
a2 <- df_dat |> 
    as_tibble() |> 
    group_by(country) |> 
    dplyr::summarize(
        EF = sum(d_EF, na.rm = TRUE),
        Gini = sum(d_gini, na.rm = TRUE),
        GNI = sum(d_gni, na.rm = TRUE)
    ) |> 
    mutate(Group = as.character(stab@clusterObjs$kmeans$`3`$cluster)) |> 
    left_join(
        dat |> select(country, country_code) |> 
            filter(!is.na(country_code)) |> unique()) |> 
    ggplot(aes(GNI, Gini)) +
    geom_segment(
        aes(x = 0, y = 0, xend = GNI, yend = Gini, 
            color = as.factor(Group)), show.legend = FALSE,
        arrow = arrow(length = unit(0.1, "cm")), size = 0.15) +
    geom_text(aes(x = GNI * 1.1, y = Gini *1.1, 
                  label = country_code, colour = Group), 
              size = 1, show.legend = FALSE) +
    scale_color_manual("Groups", values = c("#73B3A3","#FEA621","#5BA4CA")) +
    theme_light(base_size = 6) + labs(x = "GNI [log]") +
    theme(panel.grid = element_blank(), legend.position = c(0.8,0.2),
          legend.text = element_text(size = 4), legend.title = element_text(size = 4),
          legend.key.width = unit(2,"mm"), legend.key.height = unit(2, "mm"))

f2 <- c2 + annotation_custom(
        grob = ggplotGrob(a2),
        xmin = 0.8, xmax = 2.5, ymin = 40, ymax = 62
    )

# ggsave(
#     filename = "Gini_GNI.png", path = "figures/", device = "png",
#     width = 4, height = 4, dpi = 300, bg = "white", 
#     plot = f2
# )

```{r fig.height=3, fig.width=3}
a3 <- df_dat |> 
    as_tibble() |> 
    group_by(country) |> 
    dplyr::summarize(
        EF = sum(d_EF, na.rm = TRUE),
        Gini = sum(d_gini, na.rm = TRUE),
        GNI = sum(d_gni, na.rm = TRUE)
    ) |> 
    mutate(Group = as.character(stab@clusterObjs$kmeans$`3`$cluster)) |> 
    left_join(
        dat |> select(country, country_code) |> 
            filter(!is.na(country_code)) |> unique()) |> 
    ggplot(aes(GNI, EF)) +
    geom_segment(
        aes(x = 0, y = 0, xend = GNI, yend = EF, 
            color = as.factor(Group)), show.legend = FALSE, 
        arrow = arrow(length = unit(0.1, "cm")), size = 0.15) +
    geom_text(aes(x = GNI * 1.1, y = EF *1.1, 
                  label = country_code, colour = Group), 
              size = 1, show.legend = FALSE) +
    scale_color_manual("Groups", values = c("#73B3A3","#FEA621","#5BA4CA")) +
    theme_light(base_size = 6) + labs(y = "Ecological footprint") +
    theme(panel.grid = element_blank(), legend.position = c(0.8,0.8),
          legend.text = element_text(size = 4), legend.title = element_text(size = 4),
          legend.key.width = unit(2,"mm"), legend.key.height = unit(2, "mm"))

f3 <- c3 + annotation_custom(
        grob = ggplotGrob(a3),
        xmin = 0.8, xmax = 2.8, ymin = 8, ymax = 18
    )

# ggsave(
#     filename = "EF_GNI.png", path = "figures/", device = "png",
#     width = 4, height = 4, dpi = 300, bg = "white", 
#     plot = f3
# )

# ggsave(
#     filename = "trilema_ensamble.png", device = "png", path = "figures/",
#     width = 7.5, height = 3, bg = "white", dpi = 300, 
#     plot = f1 + labs(tag = "A") + f2 + labs(tag = "B") +  f3 + labs(tag = "C") 
# )
```

### 2022-03-02 inequality-SDGs.Rmd
data explorations for SDGs datasets

key |> pull(topic) |> unique()
```

>>>>

key |> pull(topic) |> unique()

countries <- read_csv(
    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGCountry.csv") |> 
    janitor::clean_names()

countries_list <- countries |> filter(!is.na(currency_unit)) |> pull(country_code)

```

```{r}
dat$indicator_name |> unique() |> length()
```
Extract environmental variables only

```{r}
env_topics <- key$topic |> str_subset("Environment")

env_vars <- key |> 
    filter(topic %in% env_topics) |> 
    pull(indicator_name) |> 
    unique()
```

```{r}
n_countries <- dat |> filter(country_code %in% countries_list) |> 
    pull(country_code) |> unique() |> length()

p <- dat |>
    filter(indicator_name %in% env_vars, country_code %in% countries_list) |> 
    select(-x35) |> 
    pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |> 
    mutate(year = str_remove(year, "x"), year = as.numeric(year)) |> 
    group_by(indicator_name, year) |> 
    mutate(nas = is.na(value)) |> 
    summarise(missing = sum(nas) / n_countries) |> 
    ggplot(aes(year, indicator_name)) +
    geom_tile(aes(fill = missing)) +
    scale_fill_viridis_c() +
    theme_light(base_size = 6)

ggsave(
    plot = p, filename = "WB_environment_SDGs.png", path = "figures/", device = "png",
    width = 6, height = 4, dpi = 400, bg = "white"
)

p + theme_light(base_size = 8)
```
And the inequality variables

```{r}
ineq_topics <- c("Financial Sector: Access", )

key$topic |> unique()
```

## To-dos

Wait until Maike and Tong revise the selection of variables. Take the analysis from there. Alternatively, select a couple of topics as example, just to explore how to deal with missing values and what is doable. Once one take missing values into account, the number of variables and countries get reduced quickly.
